### 2015-09-14 Manuscript/SW_early_life_telomeres_SOM_v9.Rmd
Updated figures and manuscript - standardised all response variables for within year analyses. Added markdown docx template (myref.docx)

@@

#Set wd one level up
setwd('..')

#Load and clean data

source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = getwd())
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 4,
               fig.height= 8,
               fig.show='asis',
               dev='png')

@@

FigS2 <- ggplot(ddFigS2,
       aes(x = Ageclass,
           y = TROCmean)) +
  geom_pointrange(aes(min=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  ylab('Corrected telomere loss') + 
  xlab('') + 
  theme_classic() +
  ylim(0,0.002) +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

@@

**Figure S2** Telomere loss (mean $\pm$ s.e.) in juvenile Seychelles warblers. Telomere loss a standardised and corrected value (see methods for details) based on rates of lost between the sampling event of birds as a juvnile, and the first available sampling as an adult. Values above zero indicate telomere shortening in early life.

```{r Figure S3,fig.width=4,fig.height=4}

temp <- FlSA

ddFigS3 <- ddply(temp,
                 .(LayYear),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

ddFigS3 <- ddFigS3[complete.cases(ddFigS3),]
ddFigS3$LayYear <- as.numeric(paste(ddFigS3$LayYear))

FigS3 <- ggplot(ddFigS3,
       aes(x = LayYear,
           y = TLKBmean)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  ylab('Telomere length (kb)') + 
  xlab('Lay Year') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

FigS3

@@

**Figure S3** Cross-sectional juvenile telomere length (mean $\pm$ s.e.) across birth years in Seychelles warblers

>>>>

```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 4,
               fig.height= 8,
               fig.show='asis',
               dev=c('png','pdf'))

@@

#Load and clean data
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')

@@

FigS2 <- ggplot(ddFigS2,
       aes(x = Ageclass,
           y = TROCmean)) +
  geom_pointrange(aes(min=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  ylab('Corrected telomere loss') + 
  xlab('') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

@@

**Figure S2** Longitudinal telomere loss (mean $\pm$ s.e.) in juvenile Seychelles warblers. Telomere loss a standardised and corrected value (see methods for details) based on rates of lost between the sampling event of birds as a juvnile, and the first available sampling as an adult. Values above zero indicate telomere shortening in early life.

```{r Figure S3,fig.width=5,fig.height=9}

###########
#Fig S3A

temp <- subset(FlSAall,Ageclass == 'FL')

#Run the model
lm3 <- lmer(TLKB ~ TQspace + Insect + Sex + Tarsus + BodyMass + (1|LayYear),
            data = temp,REML=FALSE)

lm3 <- standardize(lm3)

#Model averaging
lm4 <- dredge(lm3)
lmavg <- model.avg(lm4,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table1 <- imptable(lmavg,myimp)
table1 <- signif(table1,2)

#Sort out row/col names
nams <- rownames(table1)
nams[nams == 'c.Ageclass'] <- 'Age class'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.Insect'] <- 'Food av.'
nams[nams == 'z.TQspace'] <- 'TQ'
nams[nams == 'z.Helper'] <- 'Helpers'

rownames(table1) <- nams
colnames(table1) <- c('Estimate','SE','2.5%','97.5%','RI')

table1 <- data.frame(table1,myx=row.names(table1))

FigS3A <- forestplot(table1) +
  annotate('text',x=0.75,y=3,label='A')

###########
#Fig S3B

temp <- subset(FlSAall,Ageclass == 'SA')

#Run the model
lm3 <- lmer(TLKB ~ TQspace + Insect + Sex + Tarsus + BodyMass + (1|LayYear),
            data = temp,REML=FALSE)

#lm3 <- standardize(lm3)

#Model averaging
lm4 <- dredge(lm3)
lmavg <- model.avg(lm4,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table1 <- imptable(lmavg,myimp)
table1 <- signif(table1,2)

#Sort out row/col names
nams <- rownames(table1)
nams[nams == 'c.Ageclass'] <- 'Age class'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.Insect'] <- 'Food av.'
nams[nams == 'z.TQspace'] <- 'TQ'
nams[nams == 'z.Helper'] <- 'Helpers'

rownames(table1) <- nams
colnames(table1) <- c('Estimate','SE','2.5%','97.5%','RI')

table1 <- data.frame(table1,myx=row.names(table1))

FigS3B <- forestplot(table1) +
  annotate('text',x=0.75,y=3,label='B')

multiplot(FigS3A,FigS3B)

@@

**Figure S3** Factors affecting telomere length in Seychelles warbler fledglings **(A)** and subadults **(B)**. Points and error bars represent model averaged estimates and 95% confidence intervals for all explanatory terms used in a linear model with telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; Food av. = annual food availability, TQ = territory quality).

### 2015-09-14 Manuscript/SW_early_life_telomeres_tables_and_figures v9.Rmd
Updated figures and manuscript - standardised all response variables for within year analyses. Added markdown docx template (myref.docx)

---
output: word_document
---
```{r setwd, include = F}

#Set working directory up one level
setwd('..')

@@

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = getwd())
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 4,
               fig.height= 8,
               fig.show='asis',
               dev='png')

@@

**Figure 1** Factors affecting telomere length in juvenile Seychelles warblers. **A** Model averaged estimates and 95% confidence intervals (points and error bars, respectively) for all explanatory terms used in a linear model with juvenile telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; Food av = annual food availability, TQ = territory quality). **B** Annual food availability and **C** body condition in relation to telomere length. Points and error bars in **A** are mean $\pm$ 95% confidence interval telomere lengths for each year or group size (means are used for plotting purposes only), and the lines represent fitted values and from linear regressions.

**Figure 2** Juvenile telomere length and lifespan in Seychelles warblers. **A** Model averaged estimates and 95% confidence intervals (points and error bars, respectively) for all explanatory terms used in a generalised linear model with a poisson error structure and lifespan as the response variable. Only juveniles that survived their first year of life were included in the model. Numbers in brackets are the relative importance of each term in the top model set (see main text for details). Food av = annual food availability, TL = telomere length TQ = territory quality. **B** Raw telomere length and annual food availability in relation to lifespan. Annual food availability was modeled as a covariate (see main text for details) but is plotted as a factor here for visualisation purposes. The lines represent fitted values from a Poisson regression.

@@

```{r Figure 1,fig.width=8,fig.height=9}

###########
#Fig 1A

#Run the model
lm1 <- lmer(TLKB~ TQspace + Insect + GroupSize + Sex + Condition + Helper + (1|LayYear),
            data = FlSAall,REML=FALSE)

lm1 <- standardize(lm1)

#Model averaging
lm2 <- dredge(lm1)
lmavg <- model.avg(lm2,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table1 <- imptable(lmavg,myimp)
table1 <- signif(table1,2)

#Sort out row/col names
nams <- rownames(table1)
nams[nams == 'c.Ageclass'] <- 'Age class'
nams[nams == 'z.Condition'] <- 'Body condition'
nams[nams == 'z.GroupSize'] <- 'Local density'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.Insect'] <- 'Food av.'
nams[nams == 'z.TQspace'] <- 'TQ'
nams[nams == 'z.Helper'] <- 'Helpers'

rownames(table1) <- nams
colnames(table1) <- c('Estimate','SE','2.5%','97.5%','RI')

table1 <- data.frame(table1,myx=row.names(table1))

Fig1A <- forestplot(table1) +
  annotate('text',x=0.5,y=4,label='A')

######################################
temp <- FlSAall
temp$Insect <- factor(temp$Insect)

ddFig1B <- ddply(temp,
                 .(Insect),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = 1.96*se(TLKB))
ddFig1B <- ddFig1B[complete.cases(ddFig1B),]
ddFig1B$Insect <- as.numeric(paste(ddFig1B$Insect))

Fig1B <- ggplot(ddFig1B,
       aes(x = Insect,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('Annual food availability') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  stat_smooth(data=FlSAall,aes(x=Insect,y=TLKB),method=lm,col='black',se=F) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  scale_colour_manual(values = wes_palette("Royal1")) +
  annotate('text',x=8,y=12,label='B')

Fig1C <- ggplot(FlSAall,
                aes(x = Condition,
                    y = TLKB)) +
    geom_point() +
  ylab('Telomere length (kb)') + 
  xlab('Body Condition') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  stat_smooth(method=lm,col='black',se=F) +
  scale_colour_manual(values = wes_palette("Royal1")) +
  annotate('text',x=-2.5,y=15,label='C')

multiplot(Fig1A,Fig1B,Fig1C,layout=matrix(c(1,1,1,1,1,1,2,3,2,3), nrow=5, byrow=TRUE))

@@

```{r Figure 2}

t1 <- data.frame(with(FlSAall,flexsurvreg(Surv(Lifespan+1,Died)~Condition+TLKB+Insect, dist="lnorm"))$res)
t1 <- t1[3:5,]
rownames(t1) <- c('Body condition', 'Tel. length', 'Food av.')

Fig2A <- ggplot(t1,aes(x = rownames(t1),y = est)) +
  geom_point() + 
  geom_pointrange(aes(ymin = L95.,ymax = U95.)) +
    theme_classic() +
  ylab('Estimate') + 
  xlab('') + 
  theme_classic() +
  theme(axis.text.x = element_text(angle=90,size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  geom_abline(intercept = 0,slope = 0,lty=2)

s1 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = FlSA)
Fig2B <- ggsurv(s1,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion Survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  theme(legend.title = element_blank())

multiplot(Fig2A,Fig2B)

@@

>>>>

---
output:
  word_document:
    reference_docx: myref.docx
---
```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 4,
               fig.height= 8,
               fig.show='asis',
               dev=c('png','pdf'))

```{r load and clean, include = F}

@@

@@

**Figure 1** Telomere length and age Seychelles warblers, based on cross-sectional **(A)** and longitudinal **(B)** data. Points and error bars represent mean and 95% confidence intervals, respectively. In **B**, telomere loss is a standardised and corrected rate (see methods for details) based on differences in telomere length between the sampling event of birds as a juvnile, and the first available sampling as an adult. Values above zero indicate telomere shortening in early life.

**Figure 2** Temporal variation in early-life telomere length in the Seychelles warbler based on cross-sectional **(A,B)** and longitudinal **(C,D)** data (see Figure 1 and main text for details). Points and error bars represent mean and standard error telomere length for each summer (blue) or winter (red) field season. In **A** and **C** points are ordered temporally, while in **B** and **D** telomere length/loss for each season is plotted against island-wide food availability. Lines and shaded areas represent fitted values and 95% confidence limits from a linear regression.

**Figure 3** Factors affecting telomere length in Seychelles warbler chicks. **A** Model averaged estimates and 95% confidence intervals for all explanatory terms used in a linear model with telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; Food av. = annual food availability, TQ = territory quality). **B** Presence/absence of helpers, **C** sex and **D** tarsus length in relation to raw telomere length. Points and error bars in **B and C** are mean $\pm$ standard error telomere lengths for each group, and the line and shaded areas from **D** represent fitted values and 95% confidence limits from a linear regression.

**Figure 4** Factors affecting telomere length in Seychelles warblers fledglings **(A)** and subadults **(B)**. Model averaged estimates and 95% confidence intervals for all explanatory terms used in a linear model with  telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; Food av. = annual food availability, TQ = territory quality).

**Figure 5** Kaplan-Meier curves showing the relationship between telomere length and survival in chick (**A**) and juvenile (**B**) Seychelles warblers. Telomere length is binned into groups here for visualisation purposes only (long and short = greater than or less than median telomere length, repectively).

@@

```{r Figure 1,fig.width=8,fig.height=9}
######################
#1A
######################

temp <- dd
levels(temp$Ageclass) <- c('Chicks',rep('Fledglings',3),'Sub-adults','Adults')

ddFig1A <- ddply(temp,
                 .(Ageclass),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

Fig1A <- ggplot(ddFig1A,
       aes(x = Ageclass,
           y = TLKBmean)) +
  ylab('Corrected telomere loss') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'A',x=0.5,y=6.2)

######################
#1B
######################

temp <- Loss
levels(temp$Ageclass) <- c('Chicks','Fledglings','Sub-adults')

ddFig1B <- ddply(temp,
                 .(Ageclass),
                 summarize,
                 TROCmean = mean(TROC),
                 TROCse = se(TROC))

Fig1B <- ggplot(ddFig1B,
       aes(x = Ageclass,
           y = TROCmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  annotate('text',label = 'B',x=0.5,y=0.0018)

multiplot(Fig1A,Fig1B)

@@

```{r Figure 2,fig.width=9,fig.height=9}

ddFig2A <- ddply(juv,
                 .(CatchYear,Season),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))
ddFig2A <- ddFig2A[complete.cases(ddFig2A),]
ddFig2A <- subset(ddFig2A,TLKBse < 4)

Fig2A <- ggplot(ddFig2A,
       aes(x = CatchYear,
           y = TLKBmean,
           col = Season)) +
  ylab('Telomere length') + 
  xlab('Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse),
                  position = position_dodge(width = 1)) +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'A',x=1995,y=10)

####################

temp <- juv
temp$Insect <- factor(temp$Insect)

ddFig2B <- ddply(temp,
                 .(Insect,Season),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))
ddFig2B$Insect <- as.numeric(paste(ddFig2B$Insect))
ddFig2B <- ddFig2B[complete.cases(ddFig2B),]

Fig2B <- ggplot(ddFig2B,
       aes(x = Insect,
           y = TLKBmean,
           col = Season)) +
  ylab('Telomere length') + 
  xlab('Food availability') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  stat_smooth(method = 'lm',se=T,col = 'darkgrey') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'B',x=-1.5,y=8)

################################

ddFig2C <- ddply(Loss,
                 .(CatchYear,Season),
                 summarize,
                 TROCmean = mean(TROC),
                 TROCse = se(TROC))
ddFig2C <- ddFig2C[complete.cases(ddFig2C),]
ddFig2C <- subset(ddFig2C,TROCse < 4)

Fig2C <- ggplot(ddFig2C,
       aes(x = CatchYear,
           y = TROCmean,
           col = Season)) +
  ylab('Corrected telomere loss') + 
  xlab('Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TROCmean-TROCse,ymax=TROCmean+TROCse),
                  position = position_dodge(width = 1)) +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'C',x=1995,y=0.009)

####################################
temp <- Loss
temp$Insect <- factor(temp$Insect)

ddFig2D <- ddply(temp,
                 .(Insect,Season),
                 summarise,
                 TROCmean = mean(TROC),
                 TROCse = se(TROC))
ddFig2D$Insect <- as.numeric(paste(ddFig2D$Insect))
ddFig2D <- ddFig2D[complete.cases(ddFig2D),]

Fig2D <- ggplot(ddFig2D,
       aes(x = Insect,
           y = TROCmean,
           col = Season)) +
  geom_pointrange(aes(min=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  ylab('Corrected telomere loss') + 
  xlab('Food availability') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  stat_smooth(method = 'lm', se=T, col = 'darkgrey') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'D',x=-1.5,y=0.0075)

multiplot(Fig2A,Fig2B,Fig2C,Fig2D,layout=matrix(1:4,2,2))

@@

**Figure 3**

```{r Figure 3,fig.width=8,fig.height=9}

######################
#3A
######################

#Run the model
lm2 <- lm(cenTL ~ cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ,
            data = chickall)

#Model averaging

lm3 <- dredge(lm2)
lmavg <- model.avg(lm3,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'

rownames(table2) <- nams
colnames(table2) <- c('Estimate','SE','2.5%','97.5%','RI')

table2 <- data.frame(table2,myx=row.names(table2))

Fig3A <- forestplot(table2) +
  annotate('text',label = 'A',x=0.6,y=1.2)

######################
#3B
######################

temp <- chicks

ddFig3B <- ddply(temp,
                 .(Sex),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

Fig3B <- ggplot(ddFig3B,
       aes(x = Sex,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'B',x=0.5,y=7)

######################
#3C
######################

temp <- chicks
temp$HelperF <- ifelse(chicks$Helper>0,'Helpers present','Helpers absent')

ddFig3C <- ddply(temp,
                 .(HelperF),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

Fig3C <- ggplot(ddFig3C,
       aes(x = HelperF,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'C',x=0.5,y=8)

######################
#3D
######################

Fig3D <- ggplot(chicks,
                aes(x = Tarsus,
                    y = TLKB)) +
  ylab('Telomere length (kb)') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  geom_point(size=3) +
  stat_smooth(method = 'lm',col = 'black') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'D',x=19,y=15)

multiplot(Fig3A,Fig3B,Fig3C,Fig3D,layout=matrix(c(1:4), nrow=2, byrow=FALSE))

**Figure 4**

```{r Figure 4,fig.width=5,fig.height=9}

######################
#3A
######################
Fl <- subset(FlSAall,Ageclass == 'FL')

#Run the model
lm2 <- lm(cenTL ~ cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ,
            data = Fl)

#Model averaging
lm3 <- dredge(lm2)
lmavg <- model.avg(lm3,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)

nams <- rownames(table2)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'

rownames(table2) <- nams
colnames(table2) <- c('Estimate','SE','2.5%','97.5%','RI')

table2 <- data.frame(table2,myx=row.names(table2))

Fig4A <- forestplot(table2) +
  annotate('text',label = 'A',x=0.6,y=0.8)

######################
#4B
######################
SA <- subset(FlSAall,Ageclass == 'SA')

#Run the model
lm2 <- lm(cenTL ~ cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ,
            data = SA)

#Model averaging
lm4 <- dredge(lm2)
lmavg <- model.avg(lm4,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)

nams <- rownames(table2)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'

rownames(table2) <- nams
colnames(table2) <- c('Estimate','SE','2.5%','97.5%','RI')

table2 <- data.frame(table2,myx=row.names(table2))

Fig4B <- forestplot(table2) +
  annotate('text',label = 'B',x=0.6,y=0.6)

multiplot(Fig4A,Fig4B)

**Figure 5**

```{r Figure 5,fig.height=8,fig.width=5}

s1 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = chicks)
Fig5A <- ggsurv(s1,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='A')

###########################################

s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = FlSAall)
Fig5B <- ggsurv(s2,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='B')

multiplot(Fig5A,Fig5B)
